---
title: "Homework 2"
author: "Frederick J. Boehm"
date: "9/21/2020"
output: pdf_document
---

# Exercise 1

## i $\iff$ iii

we want to show that i $\iff$ iii. 

We follow the hint. 

$$
f(x) = f(x_0) + f'(x_0) (x - x_0) + \frac{1}{2}f''(\eta)(x - \eta) ^ 2
$$

We then see that $f''(\eta) \ge 0$ means that the last term in the right-hand side is non-negative. Thus, $f(x) \ge f(x_0) + f'(x_0) (x - x_0)$. For the reverse implication, reverse the above steps. 

We thus see that i $\iff$ iii. 

## Plot for ii


We let $f(x) = x ^ 2$ with domain $[0, 1]$. We vary $\lambda$ from zero to one below.

```{r}
library(magrittr)
library(ggplot2)
tibble::tibble(x = 1:100 / 100, xsq = x ^ 2)  %>%
  ggplot() + geom_line(aes(y = x, x = x)) + geom_line(aes(y = xsq, x = x))
```
For x = 0, y = 1, we see that 

$$f(\lambda x + (1 - \lambda)y) \le \lambda f(x) + (1-\lambda)f(y)$$

## Plot for iii

we set $x_0 = 1$ and vary $x$.


```{r}
tibble::tibble(x = 1:100 / 100, lhs = x ^ 2, rhs = 1 ^ 2 + 2 * (x - 1) )  %>%
  ggplot() + geom_line(aes(y = lhs, x = x, colour = "lhs")) + geom_line(aes(y = rhs, x = x, colour = "rhs"))

```

We plot along the horizontal axis the values of $x$ and along the 
vertical axis the values of $f(x_0) + f'(x_0) (x - x_0)$ and those of $f(x)$. "lhs" denotes $f(x)$, while "rhs" denotes $f(x_0) + f'(x_0) (x - x_0)$




# Exercise 2

## Discrete random variables

A discrete random variable $X$ taking countably many values $x_i$ has expectation $\mathbb{E}X = \sum_{i = 1} ^ {\infty}p_ix_i$. 

$$
\begin{aligned}
f(\mathbb{E}X) = f(\sum_i p_ix_i) \le \sum_i p_if(x_i) = \mathbb{E}\left(f(X)\right)
\end{aligned}
$$
where we use induction on the natural numbers to argue that the 
inequality ii from question 1 extends to countably many summands. 



## General random variables

We follow the hint.

$$
\begin{aligned}
f(X) &\ge f(\mathbb{E}X) + f'(\mathbb{E}X)\left(\left[ X - \mathbb{E}X\right]\right) \\
&\implies \mathbb{E}\left( f(X)\right) \ge f(\mathbb{E}X) + f'(\mathbb{E}X)\left(\mathbb{E}\left[ X - \mathbb{E}X\right]\right) \\
&\implies \mathbb{E}\left( f(X)\right) \ge f(\mathbb{E}X)
\end{aligned}
$$

Note that the last inequality follows from line 2 because $\mathbb{E}\left[ X - \mathbb{E}X\right] = \mathbb{E}X - \mathbb{E}X = 0$.


# Exercise 3

## Part i

We recall the definition of $f$ continuous: $f$ continuous if for any 
$\epsilon > 0$ there is a $\delta > 0$ such that 
$d_X(x,y) < \delta \implies d_Y(f(x) - f(y)) < \epsilon$. 
To see that $f$ is continuous, let $\epsilon > 0$. 
Then, observe that $d_X(x,y) < \frac{\epsilon}{L} \implies d_Y(f(x), f(y))$, 
so $\frac{\epsilon}{L}$ is the desired $
delta$.

## Part ii


## Part iii

$f$ is not differentiable because $\lim_{x \downarrow 0} f'(x) = 1$ while $\lim_{x \uparrow 0} f'(x) = -1$.

To show that $f$ is 1-lipchitz on $\mathbb{R}$, consider the arbitrary interval $(a,b)$. If $a \ge 0$ then $f(b) - f(a) = b-a$. If $b \le 0$, then $f(b) - f(a) = - (b - a)$. If $a < 0$ and $b > 0$ then $f(b) - f(a) = | b - |a|| \le |b - a|$. 


## Part iv




## Part v

No. On $\mathbb{R}$, the slope, $f'(x) = 2x$ goes to infinity as $x \to \infty$. 

To be precise, suppose that there is a constant $C$ such that $f$ is C-lipschitz. $f$ C-lipschitz on $\mathbb{R}$ means that for any two points x and y, $|f(x) - f(y)| \le C|x - y|$. 

Choose $x = C$ and $y = C + 1$. Then $f(x) = C^2$ and $f(y) = C ^ 2 + 2C + 1$, so $|f(x) - f(y)| = 2C + 1$, which is a contradiction. Thus, it must be that $f$ is not L-lipschitz on $\mathbb{R}$.








# Exercise 4




# Exercise 5




# Exercise 6


