---
title: "Homework 2"
author: "Frederick J. Boehm"
date: "9/21/2020"
output: pdf_document
---

# Exercise 1

## i $\iff$ iii

we want to show that i $\iff$ iii. 

We follow the hint. 

$$
f(x) = f(x_0) + f'(x_0) (x - x_0) + \frac{1}{2}f''(\eta)(x - \eta) ^ 2
$$

We then see that $f''(\eta) \ge 0$ means that the last term in the right-hand side is non-negative. Thus, $f(x) \ge f(x_0) + f'(x_0) (x - x_0)$. For the reverse implication, reverse the above steps. 

We thus see that i $\iff$ iii. 

## Plot for ii


We let $f(x) = x ^ 2$ with domain $[0, 1]$. We vary $\lambda$ from zero to one below.

```{r}
library(magrittr)
library(ggplot2)
tibble::tibble(x = 1:100 / 100, xsq = x ^ 2)  %>%
  ggplot() + geom_line(aes(y = x, x = x)) + geom_line(aes(y = xsq, x = x))
```
For x = 0, y = 1, we see that 

$$f(\lambda x + (1 - \lambda)y) \le \lambda f(x) + (1-\lambda)f(y)$$

## Plot for iii

we set $x_0 = 1$ and vary $x$.


```{r}
tibble::tibble(x = 1:100 / 100, lhs = x ^ 2, rhs = 1 ^ 2 + 2 * (x - 1) )  %>%
  ggplot() + geom_line(aes(x = lhs, y = rhs)) + geom_line(aes(x = lhs, y = lhs))

```

We plot along the horizontal axis the values of $f(x)$ and along the 
vertical axis the values of $f(x_0) + f'(x_0) (x - x_0)$. For reference, we 
include a line through the origin with slope 1. 





# Exercise 2

## Discrete random variables

A discrete random variable $X$ taking countably many values $x_i$ has expectation $\mathbb{E}X = \sum_{i = 1} ^ {\infty}p_ix_i$. 

$$
\begin{aligned}
f(\mathbb{E}X) = f(\sum_i p_ix_i) \le \sum_i p_if(x_i) = \mathbb{E}\left(f(X)\right)
\end{aligned}
$$
where we use induction on the natural numbers to argue that the 
inequality ii from question 1 extends to countably many summands. 



## General random variables

We follow the hint.

$$
\begin{aligned}
f(X) &\ge f(\mathbb{E}X) + f'(\mathbb{E}X)\left(\left[ X - \mathbb{E}X\right]\right) \\
&\implies \mathbb{E}\left( f(X)\right) \ge f(\mathbb{E}X) + f'(\mathbb{E}X)\left(\mathbb{E}\left[ X - \mathbb{E}X\right]\right) \\
&\implies \mathbb{E}\left( f(X)\right) \ge f(\mathbb{E}X)
\end{aligned}
$$

Note that the last inequality follows from line 2 because $\mathbb{E}\left[ X - \mathbb{E}X\right] = \mathbb{E}X - \mathbb{E}X = 0$.


# Exercise 3





# Exercise 4




# Exercise 5




# Exercise 6


