---
title: "Homework 6"
author: "Fred Boehm"
date: "12/2/2020"
output: pdf_document
classoption: landscape
---

## Part A.i

### Compute the covariance matrix of Z

$$
Var(Z_1) = \frac{1}{2}Var(X_1 + X_2) = \frac{1}{2}\left( \sigma_1 ^ 2 + \sigma_2 ^ 2\right)
$$
$$
Var(Z_2) = \frac{1}{2}Var(X_1 - X_2) = \frac{1}{2}\left( \sigma_1 ^ 2 + \sigma_2 ^ 2\right)
$$

$$
Cov(Z_1, Z_2) = \mathbb{E}(Z_1Z_2) - \mathbb{E}Z_1 \mathbb{E}Z_2 = \mathbb{E}(Z_1Z_2)
$$
and 

$$
\mathbb{E}(Z_1Z_2) = \frac{1}{2}\mathbb{E}\left( X_1 ^ 2 - X_2 ^ 2\right) = \frac{1}{2}\left( \sigma_1 ^ 2 - \sigma_2 ^ 2\right)
$$
Thus, 

$$
\Sigma = \frac{1}{2}\left(
\begin{array}{cc}
\sigma_1 ^ 2 + \sigma_2 ^ 2 & \sigma_1 ^ 2 - \sigma_2 ^ 2 \\
\sigma_1 ^ 2 - \sigma_2 ^ 2 & \sigma_1 ^ 2 + \sigma_2 ^ 2
\end{array} 
\right)
$$


### Compute $\Lambda$

The precision matrix is:

$$
\Lambda = \frac{1}{2\sigma_1^2\sigma_2^2}\left(
\begin{array}{cc}
\sigma_1 ^ 2 + \sigma_2 ^ 2 & \sigma_2 ^ 2 - \sigma_1 ^ 2 \\
\sigma_2 ^ 2 - \sigma_1 ^ 2 & \sigma_1 ^ 2 + \sigma_2 ^ 2
\end{array} 
\right)
$$



### Compute eigenvalues and eigenvectors of $\Sigma$


$$
\begin{aligned}
0 &= det\left(\frac{1}{2}
\begin{array}{cc}
a + b - \lambda & b-a \\
b-a & a + b - \lambda
\end{array} 
\right)\\
&= (\frac{a + b}{2} - \lambda)^2 - (\frac{b-a}{2})^2\\
&= ab + \lambda^2 - \lambda(a + b) = (a - \lambda)(b - \lambda)
\end{aligned}
$$
Thus, the eigenvalues of $\Sigma$ are $\sigma_1^2$ and $\sigma_2^2$.



So we now have the two eigenvalues, $\sigma_1^2$ and $\sigma_2^2$

We then seek the corresponding eigenvectors.

We see that these are $(1,1)$ and $(1, -1)$.










### Level curves of the pdf

```{r}
s1 <- 1
s2 <- 10
```

```{r}
x1 <- rnorm(n = 10000, mean = 0, sd = sqrt(s1))
x2 <- rnorm(n = 10000, mean = 0, sd = sqrt(s2))
z1 <- (x1 + x2) / sqrt(2)
z2 <- (x1 - x2) / sqrt(2)
```

```{r}
library(ggplot2)
library(magrittr)
tibble::tibble(z1 = z1, z2 = z2) %>%
  ggplot() + geom_density_2d(aes(x = z1, y = z2))
```

## Part A.ii

### Find the minimizer


### Level curves



## Part A.iii

### Find the minimizer


### Level curves


## Part B: A summary of the two articles








